Garbage Collection Evaluation Infrastructure\\ for the Cog VM

### 2018-08-20 paper.tex
take reviews into account - graph titles, add quote for 2 finger algorithm, fix typos

%% Author with single affiliation.
\author{Sophie Kaleba}
                                        %% can be repeated if necessary
\affiliation{
  %\position{Position1}
  \department{Masters student}              %% \department is recommended
  \institution{Universit\'e de Lille 1}            %% \institution is required
 % \streetaddress{Street1 Address1}
  \city{Lille}
  %\state{France}
  %\postcode{Post-Code1}
  \country{France}                    %% \country is recommended
}
\email{sophie.kaleba@etudiant.univ-lille1.fr}          %% \email is recommended

@@

Part of the business of Feenk consists in analyzing large software systems. To do so, the application parses a mse file into a model, performs some analysis, and then releases the model and generates analysis results. For business, the application analyzes closed-sourced application of companies. To build an open-source benchmark, we analyse open-source applications instead of closed-source ones.

@@

Cog's Mark-Compact algorithm, named \emph{SpurPlanningCompactor}, implements the classic planning compaction algorithm for Spur.  It uses the fact that there is room for a forwarding pointer in all objects to store the eventual position of an object in the first field. It therefore first locates a large free chunk, or Eden or a memory segment, to use as the savedFirstFieldsSpace, which it uses to store the first fields of objects that will be compacted, these fields being overwritten by each object's eventual position. It then makes at least three passes through the heap.

@@

Each pass uses a three finger algorithm, a simple extension of the classic two finger algorithm with an extra middle finger used to refer to the lowest pinned object, if any, between the to and from fingers.  Objects are moved down, starting at the first free object or chunk, provided that they fit below the lowest pinned object above the to finger.  When an object won't fit the to finger is moved above the pinned object and the third finger is reset to the next pinned object below the from finger, if any.  The scheme preserves object order amongst movable objects, making no attempt to fit objects in any gaps left below pinned objects.

@@

    \begin{subfigure}[b]{.48\textwidth}
	\includegraphics[width=\linewidth]{figures/load} 
	\caption{Load\vspace{0.2cm}}
   	\end{subfigure}\hspace{0.03\textwidth}% 
   	\begin{subfigure}[b]{.48\textwidth}
	\includegraphics[width=\linewidth]{figures/prop} 
	\caption{ExpandProperties\vspace{0.2cm}}
   	\end{subfigure}	
	\begin{subfigure}[b]{.48\textwidth}
	\includegraphics[width=\linewidth]{figures/propCache} 
	\caption{ExpandPropertiesWithCache}
	\end{subfigure}\hspace{0.03\textwidth}%
	   	\begin{subfigure}[b]{.48\textwidth}
	\includegraphics[width=\linewidth]{figures/release} 
	\caption{Release}
   	\end{subfigure}

@@

\begin{figure*}[thb]
\begin{tabular}{|c|c|l|l|l|l|l|l|}
   \hline
  & & Total   & Scavenge  & Total Full  & Compaction  & Initial  & Final \\
 Benchmark & &  Exec. time &  time & GC time &  time &  Heap size & Heap size \\
  & &  (ms) &  (ms) & (ms) &  (ms) &  (Mb) &  (Mb) \\
   \hline
   \multirow{4}{*}{Load} & Sweep 	& 159821 $\pm$ 2218 &	73599 $\pm$ 1281 	& 10863 $\pm$ 726 	&1398 $\pm$ 91 	& 193 $\pm$ 0 	& 959 $\pm$ 9.69 \\
    				    & Compact 	& 162305 $\pm$ 2790 &	72150 $\pm$ 782 	& 14091 $\pm$ 972 	&4661 $\pm$ 289 	& 193 $\pm$ 0 	& 909 $\pm$ 9.69 \\
    				    & Compact-16 	& 96043 $\pm$ 540 &	14279 $\pm$ 124 	& 12440 $\pm$ 217 	&5575 $\pm$ 36 	& 283 $\pm$ 0 	& 1055 $\pm$ 0 \\
    				    & Sweep-16 		& 91721 $\pm$ 1131 &	13862 $\pm$ 49 	& 9352 $\pm$ 73 	&1208 $\pm$ 9 	& 283 $\pm$ 0 	& 988 $\pm$ 0 \\
   \hline
   \multirow{4}{*}{ExpandProperties} 	& Sweep 		& 232940 $\pm$ 1287 &	55497 $\pm$ 238 	& 12240 $\pm$ 285 	&1178 $\pm$ 6.66 	& 959 $\pm$ 9.69 & 1888 $\pm$ 0 \\
    				    			& Compact 	& 183314 $\pm$ 13090 &	53165 $\pm$ 2343 	& 12496 $\pm$ 197 	&3775 $\pm$ 108 	& 909 $\pm$ 9.69 & 1938 $\pm$ 0 \\
    				    & Compact-16 	& 100315 $\pm$ 138 &	9372 $\pm$ 75 	& 9466 $\pm$ 87 	&3301 $\pm$ 31 	& 1055 $\pm$ 0 	& 2028 $\pm$ 0 \\
    				    & Sweep-16 		& 107914 $\pm$ 216 &	9714 $\pm$ 34 	& 9384 $\pm$ 107 	&864 $\pm$ 5 	& 988 $\pm$ 0 	& 2011 $\pm$ 0 \\
   \hline
    					& Sweep 		& 510013 $\pm$ 5812 	&	64234 $\pm$ 823 	& 23225 $\pm$ 100 	&2011 $\pm$ 10 	& 1888 $\pm$ 0 	& 1888 $\pm$ 0 \\
     	ExpandProperties	& Compact 	& 408169 $\pm$ 20444 	&	64492 $\pm$ 1234 	& 0 $\pm$ 0 		&0 $\pm$ 0 		& 1938 $\pm$ 0 	& 1938 $\pm$ 0 \\
    		WithCache	& Compact-16 	& 105389 $\pm$ 1067 	&	6467 $\pm$ 47 		& 3019 $\pm$ 24 	&937 $\pm$ 6 		& 2028 $\pm$ 0 	& 2028 $\pm$ 0 \\
    				    	& Sweep-16 	& 114325 $\pm$ 403 	&	6323 $\pm$ 23 		& 2347 $\pm$ 13 	&209 $\pm$ 6 		& 2011 $\pm$ 0 	& 2011 $\pm$ 0 \\
   \hline
   \multirow{4}{*}{Release} 	& Sweep 		& 606 $\pm$ 16 		&	0 $\pm$ 0 	& 605 $\pm$ 16 	&316 $\pm$ 12 		& 1888 $\pm$ 0 		& 534 $\pm$ 58 \\
    				    		& Compact 	& 1092 $\pm$ 2 		&	0 $\pm$ 0 	& 1092 $\pm$ 1 	&806 $\pm$ 3 			& 1938 $\pm$ 0 		& 193 $\pm$ 0 \\
    				    & Compact-16 	& 1120 $\pm$ 13 		&	0 $\pm$ 0 		& 1119 $\pm$ 13 	&829 $\pm$ 8 			& 2028 $\pm$ 0 		& 300 $\pm$ 0 \\
    				    & Sweep-16 		& 610 $\pm$ 1 			&	0 $\pm$ 0 		& 610 $\pm$ 1 		&314 $\pm$ 2 			& 2011 $\pm$ 0 		& 535 $\pm$ 203 \\
   \hline
\end{tabular} 
\caption{Moose benchmark results with standard errors.} \label{MooseRes2}
\end{figure*}

@@

\paragraph{Load.} The Load benchmark grows the memory by around 700Mb. Compaction time is smaller for the sweep algorithm, making the benchmark a little bit faster in the Mark-Sweep, which is to be expected since Sweep does not need to waste time updating pointers because objects are not moved in memory.

>>>>

%% Author with single affiliation.
\author{Sophie Kaleba}
                                        %% can be repeated if necessary
\affiliation{
  %\position{Position1}
  \department{Masters student}              %% \department is recommended
  \institution{Universit\'e Lille 1}            %% \institution is required
 % \streetaddress{Street1 Address1}
  \city{Lille}
  %\state{France}
  %\postcode{Post-Code1}
  \country{France}                    %% \country is recommended
}
\email{sophie.kaleba@etudiant.univ-lille1.fr}          %% \email is recommended

@@

Part of the business of Feenk consists in analyzing large software systems. To do so, the application parses a mse file into a model (mse is the default file format supported by Moose), performs some analysis, and then releases the model and generates analysis results. For business, the application analyzes closed-sourced application of companies. To build an open-source benchmark, we analyse open-source applications instead of closed-source ones.

@@

Cog's Mark-Compact algorithm, named \emph{SpurPlanningCompactor}, implements the classic planning compaction algorithm for Spur.  It uses the fact that there is room for a forwarding pointer in all objects to store the eventual position of an object in the first field. It therefore first locates a large free chunk, or the Eden space or a memory segment, to use as the savedFirstFieldsSpace, which it uses to store the first fields of objects that will be compacted, these fields being overwritten by each object's eventual position. It then makes at least three passes through the heap.

@@

Each pass uses a three finger algorithm, a simple extension of the classic two finger algorithm \cite{Cheney} with an extra middle finger used to refer to the lowest pinned object, if any, between the to and from fingers.  Objects are moved down, starting at the first free object or chunk, provided that they fit below the lowest pinned object above the to finger.  When an object won't fit the to finger is moved above the pinned object and the third finger is reset to the next pinned object below the from finger, if any.  The scheme preserves object order amongst movable objects, making no attempt to fit objects in any gaps left below pinned objects.

@@

    \begin{subfigure}[b]{.48\textwidth}
	\includegraphics[width=\linewidth]{figures/load} 
	\caption{LoadFromMSE\vspace{0.2cm}}
   	\end{subfigure}\hspace{0.03\textwidth}% 
   	\begin{subfigure}[b]{.48\textwidth}
	\includegraphics[width=\linewidth]{figures/prop} 
	\caption{ExpandProperties\vspace{0.2cm}}
   	\end{subfigure}	
	\begin{subfigure}[b]{.48\textwidth}
	\includegraphics[width=\linewidth]{figures/propCache} 
	\caption{ExpandPropertiesWithCache}
	\end{subfigure}\hspace{0.03\textwidth}%
	   	\begin{subfigure}[b]{.48\textwidth}
	\includegraphics[width=\linewidth]{figures/release} 
	\caption{Release}
   	\end{subfigure}

@@

\begin{figure*}[thb]
\begin{tabular}{|c|c|l|l|l|l|l|l|}
   \hline
  & & Total   & Scavenge  & Total Full  & Compaction  & Initial  & Final \\
 Benchmark & &  Exec. time &  time & GC time &  time &  Heap size & Heap size \\
  & &  (ms) &  (ms) & (ms) &  (ms) &  (Mb) &  (Mb) \\
   \hline
   \multirow{4}{*}{LoadFromMSE} & Sweep 	& 159821 $\pm$ 2218 &	73599 $\pm$ 1281 	& 10863 $\pm$ 726 	&1398 $\pm$ 91 	& 193 $\pm$ 0 	& 959 $\pm$ 9.69 \\
    				    & Compact 	& 162305 $\pm$ 2790 &	72150 $\pm$ 782 	& 14091 $\pm$ 972 	&4661 $\pm$ 289 	& 193 $\pm$ 0 	& 909 $\pm$ 9.69 \\
    				    & Compact-16 	& 96043 $\pm$ 540 &	14279 $\pm$ 124 	& 12440 $\pm$ 217 	&5575 $\pm$ 36 	& 283 $\pm$ 0 	& 1055 $\pm$ 0 \\
    				    & Sweep-16 		& 91721 $\pm$ 1131 &	13862 $\pm$ 49 	& 9352 $\pm$ 73 	&1208 $\pm$ 9 	& 283 $\pm$ 0 	& 988 $\pm$ 0 \\
   \hline
   \multirow{4}{*}{ExpandProperties} 	& Sweep 		& 232940 $\pm$ 1287 &	55497 $\pm$ 238 	& 12240 $\pm$ 285 	&1178 $\pm$ 6.66 	& 959 $\pm$ 9.69 & 1888 $\pm$ 0 \\
    				    			& Compact 	& 183314 $\pm$ 13090 &	53165 $\pm$ 2343 	& 12496 $\pm$ 197 	&3775 $\pm$ 108 	& 909 $\pm$ 9.69 & 1938 $\pm$ 0 \\
    				    & Compact-16 	& 100315 $\pm$ 138 &	9372 $\pm$ 75 	& 9466 $\pm$ 87 	&3301 $\pm$ 31 	& 1055 $\pm$ 0 	& 2028 $\pm$ 0 \\
    				    & Sweep-16 		& 107914 $\pm$ 216 &	9714 $\pm$ 34 	& 9384 $\pm$ 107 	&864 $\pm$ 5 	& 988 $\pm$ 0 	& 2011 $\pm$ 0 \\
   \hline
    					& Sweep 		& 510013 $\pm$ 5812 	&	64234 $\pm$ 823 	& 23225 $\pm$ 100 	&2011 $\pm$ 10 	& 1888 $\pm$ 0 	& 1888 $\pm$ 0 \\
     	ExpandProperties	& Compact 	& 408169 $\pm$ 20444 	&	64492 $\pm$ 1234 	& 0 $\pm$ 0 		&0 $\pm$ 0 		& 1938 $\pm$ 0 	& 1938 $\pm$ 0 \\
    		WithCache	& Compact-16 	& 105389 $\pm$ 1067 	&	6467 $\pm$ 47 		& 3019 $\pm$ 24 	&937 $\pm$ 6 		& 2028 $\pm$ 0 	& 2028 $\pm$ 0 \\
    				    	& Sweep-16 	& 114325 $\pm$ 403 	&	6323 $\pm$ 23 		& 2347 $\pm$ 13 	&209 $\pm$ 6 		& 2011 $\pm$ 0 	& 2011 $\pm$ 0 \\
   \hline
   \multirow{4}{*}{Release} 	& Sweep 		& 606 $\pm$ 16 		&	0 $\pm$ 0 	& 605 $\pm$ 16 	&316 $\pm$ 12 		& 1888 $\pm$ 0 		& 534 $\pm$ 58 \\
    				    		& Compact 	& 1092 $\pm$ 2 		&	0 $\pm$ 0 	& 1092 $\pm$ 1 	&806 $\pm$ 3 			& 1938 $\pm$ 0 		& 193 $\pm$ 0 \\
    				    & Compact-16 	& 1120 $\pm$ 13 		&	0 $\pm$ 0 		& 1119 $\pm$ 13 	&829 $\pm$ 8 			& 2028 $\pm$ 0 		& 300 $\pm$ 0 \\
    				    & Sweep-16 		& 610 $\pm$ 1 			&	0 $\pm$ 0 		& 610 $\pm$ 1 		&314 $\pm$ 2 			& 2011 $\pm$ 0 		& 535 $\pm$ 203 \\
   \hline
\end{tabular} 
\caption{Moose benchmark results with standard errors.} \label{MooseRes2}
\end{figure*}

@@

\paragraph{LoadFromMSE.} The LoadFromMSE benchmark grows the memory by around 700Mb. Compaction time is smaller for the sweep algorithm, making the benchmark a little bit faster in the Mark-Sweep, which is to be expected since Sweep does not need to waste time updating pointers because objects are not moved in memory.
